{\fontfamily{ptm}\selectfont Reading Notes}

### 2025-03-26 Notebook LaTEX/Formalized Notes/Notes 2025 03/main.tex
>w<, changes
>w<, taking notes

\date{March 2, 2020}

\maketitle
\todoinline{This paper is currently in draft mode. Check source to change options. }

\begin{abstract} 
    Reports on papers read. 
    \todo{This is just psuedo text. This is just psuedo text. This is just psuedo text. This is just psuedo text \cite{clarke_optimization_1990}. }
\end{abstract}

@@

\bibliography{references/Books.bib}

>>>>

\date{Last Compiled: \today}

\maketitle

\begin{abstract} 
    Reports on papers read. 
    This is a LaTEX file for my own notes taking. 
    It may accelerate the process of writing my thesis for my PhD degree. 
    \todo{
        This is just pseudo text. This is just pseudo text. 
        This is just pseudo text. This is just pseudo text \cite{necoara_linear_2019}. 
    }
    \todoinline{This paper is currently in draft mode. Check source to change options. }
\end{abstract}
\chapter{The Basics of Optimization Theories}

\chapter{Linear Convergence of First Order Method}
    In this chapter, we are specifically interested in characterizing linear convergence of well known first order optimization algorithms. 
    \section{Necoara's et al's Paper}
        \subsection{The Settings}
            The assumption follows give the same setting as Necoara et al. \cite{necoara_linear_2019}. 
            % NECOARA ASSUMPTIONS 
            \begin{assumption}\label{ass:necoara-2019-settings}
                Consider optimization problem: 
                \begin{align}
                    -\infty < f^+ = \min_{x \in X} f(x) . 
                \end{align}\label{problem:necoara-2019}
                $X\subseteq \RR^n$ is a closed convex set. 
                Assume projection onto $X$, denoted by $\Pi_X$ is easy. 
                Denote $X^+ = \argmin_{x \in X}f(x) \neq \emptyset$, assume it's a closed set. 
                Assume $f$ has $L_f$ Lipschitz continuous gradient, i.e: for all $x, y\in X$: 
                \begin{align*}
                    \Vert \nabla f(x) - \nabla f(y)\Vert \le L_f\Vert x - y\Vert. 
                \end{align*}
            \end{assumption}
            % BREGMAN DIV DEFINITION
            \begin{definition}[Bregman Divergence]
                Let $f:\RR^n \rightarrow \overline \RR$ be a differentiable function. 
                Define Bregman Divergence: 
                \begin{align*}
                    D_f: \RR^n \times \dom \nabla f \rightarrow \overline \RR:= 
                    (x, y) \mapsto f(x) - f(y) - \langle \nabla f(y), x - y\rangle. 
                \end{align*}
            \end{definition}
            Some immediate consequences of Assumption \ref{ass:necoara-2019-settings} now follows. 
            The variational inequality characterizing optimal solution has: 
            \begin{align*}
                x^+ \in X^+ \implies 
                (\forall x \in X)\; \langle \nabla f(x^+), x - x^+\rangle \ge 0. 
            \end{align*}
            The converse is true if $f$ is convex. 
            The gradient mapping in this case is: 
            \begin{align*}
                \mathcal G_{L_f}x = L_f(x - \Pi_{X}x). 
            \end{align*}
            % STRONG CONVEXITY DEFINITION
            \begin{definition}[Strong convexity]\label{def:necoara-scnvx}
                Suppose $f$ satisfies Assumption \ref{ass:necoara-2019-settings}. 
                Then $f \in \mathbb S(L_f, \kappa_f, X)$ is strongly convex iff 
                \begin{align*}
                    (\forall x, y\in X)\; 
                    \kappa_f \Vert x - y\Vert^2 \le 
                    D_f(x, y) \le L_f \Vert x - y\Vert^2. 
                \end{align*}
            \end{definition}
            Then it's not hard to imagine the following natural relaxation of the above conditions. 
            % DEFINITION OF WEAKER STRONG CONVEXITY
            \begin{definition}[Relaxations of Strong convexity]\label{def:necoara-weaker-scnvx}
                Suppose $f$ satisfies Assumption \ref{ass:necoara-2019-settings}. 
                Let $L_f \ge \kappa_f \ge 0$ such that for all $x \in X$, $\bar x = \Pi_{X^+} x$. 
                We define the following: 
                \begin{enumerate}
                    \item\label{def:neocara-qscnvx} Quasi-strong convexity (Q-SCNVX): $0 \le D_f(\bar x, x) - \frac{\kappa_f}{2}\Vert x - \bar x\Vert^2$. Denoeted by $\mathbb S'(L_f, \kappa_f, X)$. 
                    \item\label{def:necoara-qup} Quadratic under approximation (QUA): $0 \le D_f(x, \bar x) - \frac{\kappa_f}{2}\Vert x - \bar x\Vert^2$. Denoeted by $\mathbb U(L_f, \kappa_f, X)$. 
                    \item\label{def:necoara-qgg} Quadratic Gradient Growth (QGG): $0\le D_f(x, \bar x) + D_f(\bar x, x) - \kappa_f/2\Vert x - \bar x\Vert^2$. Denoted by $\mathbb G(L_f, \kappa_f, X)$. 
                    \item\label{def:necoara-qfg} Quadratic Function Growth (QFG): $0 \le f(x) - f^* - \kappa_f/2\Vert x - \bar x\Vert^2$. Denoted by $\mathbb F(L_f, \kappa_f, X)$. 
                    \item\label{def:necoara-eb} Error Bound (EB): $\Vert \mathcal G_{L_f}x\Vert \ge \Vert x - \bar x\Vert$. Denoted by $\mathbb E(L_f, \kappa_f, X)$. 
                \end{enumerate}
            \end{definition}

        \subsection{Major Results in the paper}
            In Necoara's et al, major results assume convexity of $f$. 
            \begin{theorem}(Q-SCNVX implies QUA)\label{thm:qscnvx-means-qua}
                Let $f$ satisfies Assumption \ref{ass:necoara-2019-settings} and assume $f$ is convex: 
                \begin{align*}
                    \mathbb S'(L_f, \kappa_f, X) \subseteq \mathbb U(L_f, \kappa_f, X). 
                \end{align*}
            \end{theorem}
            \begin{proof}
                We proof by induction. 
                Convexity of $f$ makes $X^+$ convex and $\Pi_{X^+}X$ unique for all $x \in X$. 
                Make inductive hypothesis that there exists $\kappa^{(k)} \ge 0$ such that 
                \begin{align*}
                    (\forall x \in X)\quad
                    f(x) \ge f^+ + \langle \nabla f(\Pi_{X^+}x), x - \Pi_{X^+}x\rangle 
                    + \kappa^{(k)}/2\Vert x - \Pi_{X^+}x \Vert^2. 
                \end{align*}
                The base case is true by convexity of $f$ with $\kappa_f^{(0)} = 0$. 
                Choose any $x \in X$ define $\bar x = \Pi_{X^+}x$. 
                Consider $x_\tau = \bar x + \tau(x - \bar x)$ for $\tau \in [0, 1]$. 
                Calculus rule has 
                \begin{align*}
                    f(x) &= 
                    f(\bar x) + \int_0^1 \langle \nabla f(x_\tau), x - \bar x\rangle d\tau
                    \\
                    &= 
                    f(\bar x) + \int_0^1 \tau^{-1}\langle \nabla f(x_\tau), \tau(x - \bar x)\rangle d\tau
                    \\
                    &= 
                    f(\bar x) + \int_0^1 \tau^{-1}\langle \nabla f(x_\tau), x_\tau - \bar x\rangle d\tau.
                \end{align*}
                $f$ is Q-SCNVX so
                \begin{align*}
                    f^+ - f(x_\tau) &\ge \langle \nabla f(x_\tau), \Pi_{X^+}x_\tau - x_\tau\rangle + 
                    \kappa_f/2 \Vert x_\tau - \Pi_{X^+}x_\tau\Vert^2
                    \\
                    &= 
                    \langle \nabla f(x_\tau), \bar x - x_\tau\rangle + 
                    \kappa_f/2 \Vert x_\tau - \bar x\Vert^2
                    \\
                    \iff 
                    \langle \nabla f(x_\tau), x_\tau - \bar x\rangle
                    &\ge f(x_\tau) - f^+ + \kappa_f/2\Vert x_\tau -\bar x\Vert^2. 
                \end{align*}
                We used $\Pi_{X^+}x_\tau = \bar x$ by convexity of $f$. 
                Therefore:
                {\footnotesize
                \begin{align*}
                    f(x) &\ge 
                    f(\bar x) + 
                    \int_0^1 \tau^{-1} \left(
                        f(x_\tau) - f^+ + \frac{\kappa_f}{2}\Vert x_\tau - \bar x\Vert^2
                    \right) d\tau
                    \\
                    &= 
                    f(\bar x) + 
                    \int_0^1 
                    \tau^{-1} \left(
                            f(x_\tau) - f^+ 
                        \right)
                        + \frac{\tau\kappa_f}{2}\Vert x - \bar x\Vert^2
                    d\tau
                    \\
                    &\ge 
                    f(\bar x) + 
                    \int_0^1 
                    \tau^{-1} \left(
                            \langle 
                                \nabla f(\Pi_{X^+}x_\tau), x_\tau - \Pi_{X^+}x_\tau
                            \rangle
                            + \frac{\kappa_f^{(k)}}{2} \Vert x_\tau - \Pi_{X^+}x_\tau\Vert^2
                        \right)
                        + \frac{\tau\kappa_f}{2}\Vert x - \Pi_{X^+}x_\tau\Vert^2
                    d\tau
                    \\
                    &= 
                    f(\bar x) + 
                    \int_0^1 
                    \tau^{-1} \left(
                            \langle 
                                \nabla f(\bar x), x_\tau - \bar x
                            \rangle
                            + \frac{\kappa_f^{(k)}}{2} \Vert x_\tau - \bar x\Vert^2
                        \right)
                        + \frac{\tau\kappa_f}{2}\Vert x - \bar x\Vert^2
                    d\tau
                    \\
                    &= 
                    f(\bar x) + 
                    \int_0^1 
                        \langle 
                            \nabla f(\bar x), x - \bar x
                        \rangle
                        + \frac{\tau\kappa_f^{(k)}}{2} \Vert x - \bar x\Vert^2
                        + \frac{\tau\kappa_f}{2}\Vert x - \bar x\Vert^2
                    d\tau
                    \\
                    &= 
                    f(\bar x) + 
                    \langle 
                        \nabla f(\bar x), x - \bar x
                    \rangle
                    +
                    \frac{\kappa^{(k)}_f + \kappa_f}{4}
                    \Vert x - \bar x\Vert^2. 
                \end{align*}
                }
                This is the new inductive hypothesis, and it has $\kappa_f^{(k + 1)} = (\kappa_f^{(k)} + \kappa_f)/2$. 
                The induction admits recurrence: 
                \begin{align*}
                    \kappa_f^{(n)} = (1/2^n)(\kappa_f^{(0)} + (2^n - 1)\kappa_f). 
                \end{align*}
                Inductive hypothesis is true for $\kappa_f^{(0)} = 0$ and $f$ being convex is sufficient. 
                It has $\lim_{n\rightarrow \infty} \kappa_f^{(n)} = \kappa_f$. 
            \end{proof}
            \begin{remark}
                This is Theorem 1 in the paper. 
                Convexity assumption of $f$ makes $X^+$ convex, so the projection is unique, and it has $\Pi_{X^+}x_\tau = \bar x$ for all $\tau \in [0, 1]$. 
                In addition, the inductive hypothesis has $\kappa_f^{(n)} \ge 0$, which is not sufficient for convexity, but necessary. 
                The projectin property remains true for nonconvex $X^+$, however the base case require rethinking. 
            \end{remark}
            \begin{theorem}[Q-SCNVX implies QGG]
                Under Assumption \ref{ass:necoara-2019-settings} and convexity of $f$, it has 
                \begin{align*}
                    \mathbb S'(L_f, \kappa_f, X) \subseteq \mathbb G(L_f, \kappa_f, X). 
                \end{align*}
            \end{theorem}
            \begin{proof}
                If $f \in \mathbb S'(L_f, \kappa_f, X)$ then Theorem \ref{thm:qscnvx-means-qua} has $f \in \mathbb U(L_f, \kappa_f, X)$. 
                Then, add \ref{def:necoara-qup}, \ref{def:neocara-qscnvx} in Definition \ref{def:necoara-weaker-scnvx} yield the results. 
            \end{proof}
            \begin{remark}
                This is Theorem 2 in the Necoara et al. \cite{necoara_linear_2019}, right after it claims $\mathbb U(L_f, \kappa_f, X)\subseteq \mathbb G(L_f, \kappa_f/2, X)$ under convexity. 
            \end{remark}
            \begin{theorem}[sufficiency of QFG]\label{thm:qfg-suff}

            \end{theorem}

@@

\bibliography{references/refs.bib}
